Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments

### 2016-09-26 papers/2016-Special-Issue/v-special-issue-2016.tex
numbered sections

\documentclass[%
%			% choose man or doc immediately below
doc,			% doc is single spaced
%man,		% man is double spaced
floatsintext,%
apacite%
]{apa6}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{booktabs}
\usepackage{layout}
\usepackage{apacite}
\usepackage{datetime}

@@

\title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
\date{\today}
\authornote{\today ~\currenttime}
\shorttitle{Is Vagueness Beneficial for Hearers?}
\twoauthors{Matt Green}{Kees van Deemter}
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}

@@

\begin{document}
\maketitle

\section{Introduction}\label{introduction}
Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).

@@

\subsubsection{General Methodology}\label{method}
The experiments reported in the present paper put the cost reduction hypothesis to the test. 
The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. 
Like Lipman, we focus on situations where numerical information is used in order to identify a referent. 
Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.

@@

% MG has not updated these stats yet at 26 Sept 2016
(H1) There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
The results showed that vagueness was beneficial for comparison but detrimental for matching. 
(H2) There was a main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
(H3) Vagueness exerted effects in different directions for the comparison task and for the matching task. 
Separate analyses were conducted at each level of the selection task to see whether within each task type there were significant effects of vagueness. 
There were: in the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-.07, se=.02, t=3.52, p<.01$). 
In the matching task vagueness significantly slowed response times compared with crisp controls ($\beta=-.07, se=.02, t=-2.89, p<.05$).

@@

\begin{table}[htbp]
\centering
\caption{Experiment 4: Instructions}
\label{Instructions for e4}
\begin{tabular}{llll}
vagueness	&	instruction		&	selection							&	instruction: Choose a square with \ldots\\
			&	format		&	task		&			\\
\toprule
\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& the same number of dots as the target \\ 
	& 	&	comparison										& fewer dots than the target \\
	\midrule
\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& about the same number of dots as the target \\ 
	& 	&	comparison										& far fewer dots than the target \\ 
\bottomrule
\end{tabular}
\end{table}%

@@

% KvD Something missing here?

@@

\paragraph{Implications for practical NLG systems}
Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests. The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG. Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question. Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.

@@

%\bibliography{bibliography}

\bibliography{v-special-issue-2016}
\end{document}

>>>>

\documentclass[%
%			% choose man or doc immediately below
%doc,	% doc is single spaced
man,		% man is double spaced
floatsintext,%
apacite%
]{apa6}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{booktabs}
\usepackage{layout}
\usepackage{apacite}
\usepackage{datetime}

@@

\title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
\date{\today}
\authornote{latex compiled \today ~\currenttime}
\shorttitle{Is Vagueness Beneficial for Hearers?}
\twoauthors{Matt Green}{Kees van Deemter}
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}

@@

\begin{document}
\maketitle   
\setcounter{secnumdepth}{3}
\section{Introduction}\label{introduction}
Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).

@@

\subsection{General Methodology}\label{method}
The experiments reported in the present paper put the cost reduction hypothesis to the test. 
The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. 
Like Lipman, we focus on situations where numerical information is used in order to identify a referent. 
Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.

@@

% MG has not updated these stats yet at 26 Sept 2016
(H1) There was no significant main effect of vagueness ($\beta =.003$, $se=.014$, $t=.202$, $p=.841$). 
The results showed that vagueness was beneficial for comparison but detrimental for matching. 
(H2) There was a main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
(H3) Vagueness exerted effects in different directions for the comparison task and for the matching task. 
Separate analyses were conducted at each level of the selection task to see whether within each task type there were significant effects of vagueness. 
There were: in the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-.07, se=.02, t=3.52, p<.01$). 
In the matching task vagueness significantly slowed response times compared with crisp controls ($\beta=-.07, se=.02, t=-2.89, p<.05$).

@@

\begin{table}[htbp]
\centering
\caption{Experiment 4: Instructions}
\label{Instructions for e4}
\begin{tabular}{llll}
vagueness	&	instruction		&	selection							&	instruction: Choose a square with \ldots\\
			&	format		&	task		&			\\
\toprule
\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& \ldots the same number of dots as the target \\ 
	& 	&	comparison										& \ldots fewer dots than the target \\
	\midrule
\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& \ldots about the same number of dots as the target \\ 
	& 	&	comparison										& \ldots far fewer dots than the target \\ 
\bottomrule
\end{tabular}
\end{table}%

@@

% KvD Something missing here?
\textcolor{red}{KvD Something missing here?}

@@

@@

\setcounter{secnumdepth}{0}
\bibliography{v-special-issue-2016}
\end{document}
